Windowed least-squares model reduction for dynamical systems

### 2020-10-13 main.tex
made minor edits

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) approach for model-reduction
	of dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual within a low-dimensional space--time trial
	subspace over time windows. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology recover Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection. In
	addition, the approach
	addresses key deficiencies in existing model-reduction
	techniques, e.g., the dependence of LSPG and space--time LSPG projection on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for both Galerkin and LSPG projection.  We consider two types of
	space--time trial
	subspaces within the proposed approach: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model. For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize then optimize) and indirect (i.e., optimize then discretize).
	Numerical experiments conducted using trial subspaces characterized by spatial
	dimension reduction demonstrate that the \methodAcronym\
	approach can yield more accurate solutions with lower space--time residuals than
	Galerkin and LSPG projection.

@@

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods~\cite{bochev_leastsquares,bochev_lsfem_book}.}
Researchers have developed such residual-minimizing model-reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with Galerkin projection. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
Petrov--Galerkin projection applied to the time-continuous
full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
in the limit as the time step goes to zero (i.e., Galerkin projection 
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG often yields more
accurate and stable solutions than Galerkin~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin.

@@

The objectives of this work are to overcome the shortcomings of existing
residual-minimizing methods, and to provide a unifying framework from which
existing methods can be assessed. In essence, the proposed \textit{\methodNameLower\
(\methodAcronym)} \approachKwd\ sequentially minimizes the FOM
residual over a sequence of arbitrarily defined
time windows. The method is characterized by three notable aspects.  First, the method
minimizes the \textit{time-continuous}  residual (i.e., that associated with
the full-order model ODE). By adopting a time-continuous viewpoint, the
formulation decouples the underlying temporal discretization scheme from the
residual-minimization problem, thus addressing a key deficiency of both LSPG
and space--time LSPG. Critically, time-continuous residual minimization also exposes
two different solution methods:  a \textit{discretize-then-optimize}
(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
indirect) method. Second, the method sequentially minimizes
the residual over arbitrarily defined \textit{time windows} rather than
sequentially minimizing the residual over time steps (as in LSPG)
or over the entire time domain (as in space--time residual-minimization
methods). This equips the method with additionally flexibility that enables it
to explore more fine-grained tradeoffs between computational cost and error.
Finally, \methodAcronym\ is formulated for two kinds of space--time trial
subspaces: one that associates with spatial dimension reduction (as employed by traditional spatial-projection-based methods), and one that associates with space--time dimension reduction (as employed by space--time methods).
The above attributes allow the \methodAcronym\ approach to be viewed as a
generalization of existing model-reduction methods, as 
Galerkin, LSPG, and space--time LSPG projection correspond to specific
instances of the formulation.
Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
provides a unifying framework from which these existing approaches can be derived.

@@

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with Galerkin, LSPG, and
space--time LSPG projection. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
\approachKwd. Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods. Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs.
Section~\ref{sec:numerical_experiments} presents numerical experiments.
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
function with respect to time with the $\dot\ $ operator.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical formulation}\label{sec:math}
	We begin by providing the formulation for the full-order model,
	followed by a description of standard model-reduction methods
	classified according to the type of trial subspace they employ.
	\subsection{Full-order model}\label{sec;FOM}
We consider the full-order model to be a dynamical system expressed as a
	system of ordinary differential equations (ODEs)
\begin{equation}\label{eq:FOM}
 \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
	\stateFOMIC,\qquad t \in [0,T],
\end{equation}
where  $\stateFOM: [0,T]
	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, 
$T \in \RRplus$ denotes the final time, 
 $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
	nonlinear in its first argument. For subsequent exposition, we introduce
	$\timeSpace$ to denote the set of (sufficiently smooth) real-valued functions acting on the time
	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
	\timeSpace $.  We refer to the initial value problem defined in
	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
	although the problem of interest described in the introduction corresponds
	to a parameterized dynamical system, we suppress dependence of the FOM ODE
	\eqref{eq:FOM} on such parameters for notational convenience, as this work
	focuses on devising a model-reduction approach applicable to a specific parameter instance.

@@

\subsection{Outstanding challenges}
This work seeks to overcome the limitations of existing residual-minimizing
model-reduction methods, and to provide a unifying framework from which
existing methods can be assessed. Specifically, we look to overcome the
complex dependence of LSPG on the time discretization,
exponential time growth of the error bounds for Galerkin and LSPG,
the cubic dependence of the computational cost of ST-LSPG on the number of
degrees of freedom, and the lack of ability for ST-LSPG to perform prediction
in time. We now describe the proposed \methodNameLower\ \approachKwd\ for this
purpose. 
%\end{comment}
%\input{formulation_3}

@@

\section{Acknowledgments}
The authors thank Yukiko Shimizu and Patrick Blonigan for numerous conversations from which this work benefited. 
E.\ Parish acknowledges an appointment to the Sandia National Laboratories'
John von Neumann Postdoctoral Research Fellowship in Computational Science. 
This work was partially sponsored by Sandia's Advanced Simulation and
Computing (ASC) Verification and Validation (V\&V) Project/Task
\#103723/05.30.02.  This paper describes objective technical results and
analysis. Any subjective views or opinions that might be expressed in the
paper do not necessarily represent the views of the U.S.\ Department of Energy
or the United States Government.  Sandia National Laboratories is a
multimission laboratory managed and operated by National Technology and
Engineering Solutions of Sandia, LLC., a wholly owned subsidiary of Honeywell
International, Inc., for the U.S.\  Department of Energy's National Nuclear
Security Administration under contract DE-NA-0003525.

>>>>

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) approach for model reduction
	of dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual within a low-dimensional space--time trial
	subspace over time windows. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology recover Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection. In
	addition, the approach
	addresses key deficiencies in existing model reduction
	techniques, e.g., the dependence of LSPG and space--time LSPG projection on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for both Galerkin and LSPG projection.  We consider two types of
	space--time trial
	subspaces within the proposed approach: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model. For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize then optimize) and indirect (i.e., optimize then discretize).
	Numerical experiments conducted using trial subspaces characterized by spatial
	dimension reduction demonstrate that the \methodAcronym\
	approach can yield more accurate solutions with lower space--time residuals than
	Galerkin and LSPG projection.

@@

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods~\cite{bochev_leastsquares,bochev_lsfem_book}.}
Researchers have developed such residual-minimizing model reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with Galerkin projection. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
Petrov--Galerkin projection applied to the time-continuous
full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
in the limit as the time step goes to zero (i.e., Galerkin projection 
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG often yields more
accurate and stable solutions than Galerkin~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin.

@@

The objectives of this work are to overcome the shortcomings of existing
residual-minimizing methods, and to provide a unifying framework from which
existing methods can be assessed. In essence, the proposed \textit{\methodNameLower\
(\methodAcronym)} \approachKwd\ sequentially minimizes the FOM
residual over a sequence of arbitrarily defined
time windows. The method is characterized by three notable aspects.  First, the method
minimizes the \textit{time-continuous}  residual (i.e., that associated with
the full-order model ODE). By adopting a time-continuous viewpoint, the
formulation decouples the underlying temporal discretization scheme from the
residual-minimization problem, thus addressing a key deficiency of both LSPG
and space--time LSPG. Critically, time-continuous residual minimization also exposes
two different solution methods:  a \textit{discretize-then-optimize}
(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
indirect) method. Second, the method sequentially minimizes
the residual over arbitrarily defined \textit{time windows} rather than
sequentially minimizing the residual over time steps (as in LSPG)
or over the entire time domain (as in space--time residual-minimization
methods). This equips the method with additionally flexibility that enables it
to explore more fine-grained tradeoffs between computational cost and error.
Finally, \methodAcronym\ is formulated for two kinds of space--time trial
subspaces: one that associates with spatial dimension reduction (as employed by traditional spatial-projection-based methods), and one that associates with space--time dimension reduction (as employed by space--time methods).
The above attributes allow the \methodAcronym\ approach to be viewed as a
generalization of existing model reduction methods, as 
Galerkin, LSPG, and space--time LSPG projection correspond to specific
instances of the formulation.
Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
provides a unifying framework from which these existing approaches can be derived.

@@

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with Galerkin, LSPG, and
space--time LSPG projection. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
\approachKwd. Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods. Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs.
Section~\ref{sec:numerical_experiments} presents numerical experiments.
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
function with respect to time with the $\dot\ $ operator.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical formulation}\label{sec:math}
	We begin by providing the formulation for the full-order model,
	followed by a description of standard model reduction methods
	classified according to the type of trial subspace they employ.
	\subsection{Full-order model}\label{sec;FOM}
We consider the full-order model to be a dynamical system expressed as a
	system of ordinary differential equations (ODEs)
\begin{equation}\label{eq:FOM}
 \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
	\stateFOMIC,\qquad t \in [0,T],
\end{equation}
where  $\stateFOM: [0,T]
	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, 
$T \in \RRplus$ denotes the final time, 
 $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
	nonlinear in its first argument. For subsequent exposition, we introduce
	$\timeSpace$ to denote the set of (sufficiently smooth) real-valued functions acting on the time
	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
	\timeSpace $.  We refer to the initial value problem defined in
	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
	although the problem of interest described in the introduction corresponds
	to a parameterized dynamical system, we suppress dependence of the FOM ODE
	\eqref{eq:FOM} on such parameters for notational convenience, as this work
	focuses on devising a model reduction approach applicable to a specific parameter instance.

@@

\subsection{Outstanding challenges}
This work seeks to overcome the limitations of existing residual-minimizing
model reduction methods, and to provide a unifying framework from which
existing methods can be assessed. Specifically, we look to overcome the
complex dependence of LSPG on the time discretization,
exponential time growth of the error bounds for Galerkin and LSPG,
the cubic dependence of the computational cost of ST-LSPG on the number of
degrees of freedom, and the lack of ability for ST-LSPG to perform prediction
in time. We now describe the proposed \methodNameLower\ \approachKwd\ for this
purpose. 
%\end{comment}
%\input{formulation_3}

@@

\section{Acknowledgments}
The authors thank Yukiko Shimizu, Patrick Blonigan, and Francesco Rizzi for numerous conversations from which this work benefited. 
E.\ Parish acknowledges an appointment to the Sandia National Laboratories'
John von Neumann Postdoctoral Research Fellowship in Computational Science. 
This work was partially sponsored by Sandia's Advanced Simulation and
Computing (ASC) Verification and Validation (V\&V) Project/Task
\#103723/05.30.02.  This paper describes objective technical results and
analysis. Any subjective views or opinions that might be expressed in the
paper do not necessarily represent the views of the U.S.\ Department of Energy
or the United States Government.  Sandia National Laboratories is a
multimission laboratory managed and operated by National Technology and
Engineering Solutions of Sandia, LLC., a wholly owned subsidiary of Honeywell
International, Inc., for the U.S.\  Department of Energy's National Nuclear
Security Administration under contract DE-NA-0003525.

### 2020-10-13 numerical_experiments.tex
made minor edits

\section{Numerical experiments}\label{sec:numerical_experiments}
We now analyze the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on three benchmark problems: (1) a reproductive ROM for the Sod shock tube problem (2) a reproductive ROM for a supersonic compressible flow in a cavity, and (3) a predictive ROM for the parameterized shallow water equations. In each experiment, we compare \methodAcronymROMs\ to either the LSPG, or the Galerkin and LSPG, ROMs. In the experiments, we assess (1) the impact of minimizing the residual over an arbitrarily sized time window on the solution accuracy, (2) the impact of the time step and time scheme on \methodAcronym, (3) the impact of the window size on the convergence properties of the ROMs, and (4) the impact of the window size on the cost of the ROMs. In all experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$. We note that the first two experiments are designed to test the \textit{reproductive} ability of the ROMs, while the third experiment tests the predictive capabilities of the ROM.

@@

\item \textit{Indirect method}: We consider two \methodAcronymROMs\ solved via the indirect method. The first uses the same CN discretization at at time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$. We solve the coupled two-point boundary 
problem via the forward--backward sweep method, and compute the action of the Jacobian transpose on vectors via automatic differentiation. We use parameters $\epsilon = 10^{-6}$, $\fbsmGrowth = 1.1$, and $\fbsmDecay=2$ in Algorithm~\ref{alg:st_iter}. 
\end{itemize}
\end{itemize}

@@

Next, we study the impact of the time step on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
window size of $\Delta T = 0.1$, with time steps $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
$100$, $50$, $20$, $10$, $2$, and $1$ time instances per window). We additionally 
consider LSPG ROMs leveraging the same set of time steps. To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 10^{-4}$. Figure~\ref{fig:time_step_study} shows the time-integrated $\elltwo$-error and residual norm  
of the various ROMs. We observe that the time-integrated $\elltwo$-error and residual norm 
of the \methodAcronymROMs\ decrease and converge as the time step decreases. This is in direct contrast to the LSPG ROMs, in where the time-integrated $\elltwo$-error and residual norm display a complex dependence on the time step. This result 
demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that for all time steps considered, the \methodAcronymROMs\ 
are less than 6x the cost of LSPG; this is despite the window sizes comprising up to 100 time instances. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_error_methods_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
\caption{Time-integrated residual norm of the \methodAcronymROM.} 
\label{fig:sod_error_methods_b}
\end{subfigure}
\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.} 
\label{fig:time_step_study}
\end{center}
\end{figure}

@@

\subsubsection{Results as a function of basis dimension}
Lastly, we assess the impact of the basis dimension on the performance of the various reduced-order models. We again consider S-reduction trial subspaces. We obtain the trial subspaces by evolving the FOM for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$ with the CN scheme. We execute Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}} = 2$, $\stateIntercept = \state_0$\footnote{We observed negative pressures in the projected initial conditions for trial subspaces of dimension $K=20,30,40$ if we used the offset $\stateIntercept = \bz$. Thus, for our convergence study we employed $\stateIntercept = \state_0$ so-as to avoid negative pressures in the projected initial conditions.}, and $\romdim = 10,20,30,40,50$. We investigate reduced-order models based on Galerkin projection, LSPG projection, and the WLS approach. All methods leverage the CN time scheme with $\Delta t = 0.002$. We present results for windows of size $\Delta T^n \equiv \Delta T = 0.002,0.004,0.02,0.04,0.10$ for the WLS approach. All WLS ROMs are solved via the direct approach described previously.

@@

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/pareto_error.pdf}
\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_error_converge}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/pareto_resids.pdf}
\caption{Time-integrated residual norm of the \methodAcronymROM.} 
\label{fig:sod_residual_converge}
\end{subfigure}
\caption{Pareto fronts for the time-integrate $\elltwo$-error (left) and time-integrated residual norm (right) as a function of wall-clock time.} 
\label{fig:convergence_study_pareto}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/converge_walltimes.pdf}
\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_error_converge}
\end{subfigure}
\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs\ as a function of basis dimension. Missing entries indicate that the solver did not converge.} 
\label{fig:convergence_study}
\end{center}
\end{figure}

>>>>

\section{Numerical experiments}\label{sec:numerical_experiments}
We now analyze the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on three benchmark problems: (1) a reproductive ROM for the Sod shock tube problem (2) a reproductive ROM for a supersonic compressible flow in a cavity, and (3) a predictive ROM for the parameterized shallow water equations. In each experiment, we compare \methodAcronymROMs\ to either the LSPG, or the Galerkin and LSPG, ROMs. In the experiments, we assess (1) the impact of minimizing the residual over different sized time windows on the solution accuracy, (2) the impact of the time step and time scheme on \methodAcronym, (3) the impact of the window size on the convergence properties of the ROMs, and (4) the impact of the window size on the cost of the ROMs. In all experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$. We note that the first two experiments are designed to test the \textit{reproductive} ability of the ROMs, while the third experiment tests the predictive capabilities of the ROM. The following numerical experiments are all executed on a machine with a 2.7 GHz 12-Core Intel Xeon E5 processor. All codes are written efficiently in Python, and we report timings for runs executed in serial.

@@

\item \textit{Indirect method}: We consider two \methodAcronymROMs\ solved via the indirect method. The first uses the same CN discretization at a time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$. We solve the coupled two-point boundary 
problem via the forward--backward sweep method, and compute the action of the Jacobian transpose on vectors via automatic differentiation. We use parameters $\epsilon = 10^{-6}$, $\fbsmGrowth = 1.1$, and $\fbsmDecay=2$ in Algorithm~\ref{alg:st_iter}. 
\end{itemize}
\end{itemize}

@@

Next, we study the impact of the time step on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
window size of $\Delta T = 0.1$, with time steps $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
$100$, $50$, $20$, $10$, $2$, and $1$ time instances per window). We additionally 
consider LSPG ROMs leveraging the same set of time steps. To assess time step convergence, we compare results to a new full-order model, which is as described in Section~\ref{sec:sod_fom} but uses a fine time step of $\Delta t = 10^{-4}$. Figure~\ref{fig:time_step_study} shows the time-integrated $\elltwo$-error and residual norm  
of the various ROMs. We observe that the time-integrated $\elltwo$-error and residual norm 
of the \methodAcronymROMs\ decrease and converge as the time step decreases. This is in direct contrast to the LSPG ROMs, in where the time-integrated $\elltwo$-error and residual norm display a complex dependence on the time step. This result 
demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that for all time steps considered, the \methodAcronymROMs\ 
are less than 6x the cost of LSPG; this is despite the window sizes comprising up to 100 time instances. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_time_step_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
\caption{Time-integrated residual norm of the \methodAcronymROM.} 
\label{fig:sod_time_step_b}
\end{subfigure}
\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.} 
\label{fig:time_step_study}
\end{center}
\end{figure}

@@

\subsubsection{Results as a function of basis dimension}
Lastly, we assess the impact of the basis dimension on the performance of the various reduced-order models. We again consider S-reduction trial subspaces. We obtain the trial subspaces by evolving the FOM for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$ with the CN scheme. We execute Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}} = 2$, $\stateIntercept = \state_0$\footnote{We observed negative pressures in the projected initial conditions for trial subspaces of dimension $K=20,30,40$ if we used the offset $\stateIntercept = \bz$. Thus, for our convergence study we employed $\stateIntercept = \state_0$ so-as to avoid negative pressures in the projected initial conditions.}, and $\romdim = 10,20,30,40,50$. We investigate reduced-order models based on Galerkin projection, LSPG projection, and the WLS approach. All methods leverage the CN time scheme with $\Delta t = 0.002$. We present results for windows of size $\Delta T^n \equiv \Delta T = 0.002,0.004,0.02,0.04,0.20$ for the WLS approach. All WLS ROMs are solved via the direct approach described previously.

@@

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/pareto_error.pdf}
\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
\label{fig:sod_error_converge_pareto}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/pareto_resids.pdf}
\caption{Time-integrated residual norm of the \methodAcronymROM.} 
\label{fig:sod_residual_converge_pareto}
\end{subfigure}
\caption{Pareto fronts for the time-integrate $\elltwo$-error (left) and time-integrated residual norm (right) as a function of wall-clock time.} 
\label{fig:convergence_study_pareto}
\end{center}
\end{figure}

%\begin{figure}
%\begin{center}
%\begin{subfigure}[t]{0.45\textwidth}
%\includegraphics[width=1.\linewidth]{figs/sod/converge_walltimes.pdf}
%\caption{Relative time-integrated $\elltwo$-error of the \methodAcronymROM.}
%\label{fig:sod_error_converge}
%\end{subfigure}
%\caption{Performance metrics of the Galerkin, LSPG, and \methodAcronymROMs\ as a function of basis dimension. Missing entries indicate that the solver did not converge.} 
%\label{fig:convergence_study}
%\end{center}
%\end{figure}
